---
title: "KU Undergraduate Project"
author: "Mohammad Shamim Hasan Mandal"
date: "2022-11-12"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide",warning = F,message = F)
```


Note : 

1. Open this file inside the project
2. Then run the code chucks line by line

## Required packages

```{r load_pacakges}
# install the packages; only first time
# install.packages("geodata")
# install.packages("terra")

library(geodata) # for data
library(terra)   # for image (raster) and vector (shpaefile)
```


## Directory management

We will create separate folders for "data" and "results"

```{r dir_create}
dir.create("data")
dir.create("result")
```


## Data download

We need internet connection to download the data set for the first time.
After downloading for next time on wards no need. 


How to get help: we will use geodata package, it have few functions. One of them
is gadm(), using that we can download vector (shapefile) for country(countries).

```{r get_help}
# load the help file for the gadm() function from the geodata package
# help(gadm, geodata)
```

First, download a vector (shapefile) for Bangladesh.

```{r download_shapefile}
bd= geodata::gadm(
  country = "Bangladesh",
  level = 1,           # level 1 of administrative sub-division
  path = "./data",     # save to data folder
  version = "latest",  # version 4.1
  resolution = 1       # high resolution
  )
```

View the data. To see the vector that we have gave the name "bd", just type it 
in the console. The terra package have a SpatVector class. So, the bd object
is of SpatVector. Remember this, this is important.

```{r view_data}
# print the vector information
bd

# using head print first six rows
head(bd)

# using head get detailed information.
str(bd)
```


Now make a map. Using the "bd" SpatVector, we can use plot() function to 
make a map

```{r bd_map}
plot(bd,
     axes=T,
     main="Map of Bangladesh",
     xlab="Longitude",
     ylab="Latitude",
     col="gray"
     )
```


## Get the extent of our study area (meaning whole Bangladesh)

```{r get_extent}
# using ext() function we get the longitude(x's) and latitudes (y's):
# xmin, xmax, ymin, ymax
# xmin : Minimum longitude
# xmax : Maximum longitude
# ymin : Minimum latitude
# ymax : Maximum latitude
bd_ext <- terra::ext(bd)

```



## Get satellite image

We need to download satelite image from the 

```{r get_satellite_image}
# we will use the package luna for this. First we need to download this: only first time
# install.packages('luna', repos='https://rspatial.r-universe.dev')
library(luna)

# help document on Modis data download
# help("getModis")
```


## Get modis


List of available MODIS products

```{r list_products}
# show products that start with MOD or MYD or MCD (meaning it will show any mathcing product)
product_list <- getProducts("^MOD|^MYD|^MCD") 


# explore here or through Rstudio panel
head(product_list)

# you can save an csv file and view using other software package
write.csv(product_list,file = "./data/product_list.csv",row.names = F)
```



We will download MODIS MOD09A1 Terra Surface Reflectance 8-Day Version 061
1. Data on Google Earth Engine [MOD09A1](https://developers.google.com/earth-engine/datasets/catalog/MODIS_061_MOD09A1)
2. [LPDAAC](https://lpdaac.usgs.gov/products/mod09a1v006/)

```{r download_Modis}
# Using getModis function we can explore the data
list_mod09 <- getModis(
  product = "MOD09A1",
  start_date = "2015-01-01",
  end_date = "2015-01-10",
  aoi = bd_ext,
  download = F,
  path = tempdir()
)
```


## Now check the results

```{r download_modis2}
# List of the images
list_mod09

```


Finally to download provide Earthdata login user and password:
If you do not have earthdata.nasa.gov user account: Please create one [EarthData](https://urs.earthdata.nasa.gov/home)

```{r setup_usr_passd}
# Using package usethis::edit_r_environ
#usethis::edit_r_environ() # this will open the .Reviron file

# write following in the file. Close the file. Restart Rstudio
#
# username = "Your_account"
# password = "Your_Password"
# 

# Note: For the first time. After saveing the username and password 
# save the file and restart R. Then start from beginning of the code.
```


# Download using username and password

Be careful of the dates, because if your are searching for longer period there will
be a lot of data which will take time to download. So, rather than downloading, first
explore the data setting "download=FALSE".




```{r dowload_modis3}
# Create a new directory inside data:
dir.create("./data/modis")

# Now download the data inside that folder
mod_files <- getModis(
  product = "MOD09A1",
  start_date = "2015-01-01",
  end_date = "2015-01-10",
  aoi = bd_ext,
  download = T,  # this will download
  path = "./data/modis",
  username = Sys.getenv("username"),
  password = Sys.getenv("password")
)
```

## list the files

List the files and read the first one
```{r list_read}
# make a list of files
f<-list.files(path="./data/modis",full.names = T)

# read
r<-rast(f[1])


# plot
plot(r)

```



Reading:

1. Where to get [data/(https://gisgeography.com/free-satellite-imagery-data-list/)
2. See luna pakcage code in github [getModis](https://github.com/rspatial/luna/blob/master/R/getMODIS.R)
function.
3. Read more on [rspatial.org](https://rspatial.org/)



